\chapter{System Overview}

The delivered system follows the brief's required technology stack: React on the frontend, FastAPI on the
backend, JWT authentication, and streamed AI responses. The repository is organised as a monorepo with
the following runtime pieces:

\begin{itemize}
  \item \texttt{frontend/}: React + TypeScript application for authentication, dashboard, document editing,
  AI interactions, sharing, and export controls.
  \item \texttt{backend\_fastapi/}: FastAPI service for JWT auth, document CRUD, permissions,
  share-by-link, export jobs, AI streaming, and version history.
  \item \texttt{realtime/}: Node + Yjs relay handling websocket collaboration and presence.
  \item \texttt{shared/}: reusable session-token helpers shared between FastAPI and realtime.
\end{itemize}

The final implementation keeps the earlier Express backend only as migration/reference history. It is not
the Assignment 2 runtime path. The evaluator should treat FastAPI as the authoritative backend.

\section{Delivered Feature Surface}

The baseline assignment requirements are all covered in the integrated branch:

\begin{itemize}
  \item registration, login, refresh, logout, and password reset,
  \item document CRUD with metadata,
  \item rich-text editing with autosave,
  \item version history and restore,
  \item owner/editor/viewer access control,
  \item sharing by direct grant and share-link,
  \item realtime collaboration across browser sessions,
  \item AI rewrite, summarize, and translate,
  \item streamed AI output with cancellation,
  \item AI history, feedback, and quota visibility,
  \item PDF and DOCX export,
  \item automated backend, frontend, realtime, integration, and browser tests.
\end{itemize}

Beyond the baseline, the implementation includes bonus-tier items such as cursor/selection tracking,
share-by-link with revocation, partial AI acceptance, and Playwright end-to-end coverage.
