\chapter{Appendix}

\section{Submission Artifact Inventory}

The final submission package is composed of three coordinated artifacts:

\begin{enumerate}
  \item \textbf{This final report PDF}, which documents requirements, architecture, project management,
  and proof-of-concept alignment.
  \item \textbf{The Git repository}, which contains the runnable MVP, root orchestration scripts,
  OpenAPI/contract notes, tests, Docker Compose configuration, CI workflow, and demo runbook.
  \item \textbf{The recorded presentation}, which should cover team responsibilities, key design
  decisions, and the proof-of-concept/demo flow.
\end{enumerate}

\section{Assignment Coverage Map}

Table~\ref{tab:coverage-map} maps the assignment brief's required deliverables to the final report and the
repository artifacts. This is included deliberately to reduce any ambiguity between the rubric, the
written report, and the delivered MVP.

\input{tables/rubric-coverage}

\section{Version-Control and Collaboration Evidence}

The proof-of-concept rubric explicitly evaluates whether the Git history is clean and whether there is
evidence of team collaboration. The delivered repository supports that in three ways:

\begin{itemize}
  \item feature work was kept on named branches and integrated through pull-request-style review rather
  than as one unstructured dump into \texttt{main},
  \item commits were intentionally kept small and purposeful so architectural evolution is visible in the
  history,
  \item the repository contains both implementation artifacts and editable documentation artifacts so
  architecture changes can be versioned alongside code rather than maintained separately.
\end{itemize}

This matters for the assignment because the architecture, repo organization, and PoC are meant to reflect
the same design process rather than three unrelated deliverables.

\section{Demo Checklist}

The recommended live demo sequence is:

\begin{enumerate}
  \item Run \texttt{npm run demo:reset} from the repository root.
  \item Open the frontend at \texttt{http://localhost:5173}.
  \item Sign in as \texttt{user\_1} and create a new document.
  \item Open the same document in a second browser/tab and switch the user to \texttt{user\_2}.
  \item Grant editor access from the owner account.
  \item Type in both windows and show live sync and presence.
  \item Pause briefly to show autosave returning the header to a saved state.
  \item Trigger an AI action, wait for the async job, then demonstrate full apply or partial apply.
  \item Open version history and demonstrate revert.
  \item Export the document and show Swagger/OpenAPI at \texttt{http://localhost:3000/docs/}.
\end{enumerate}

\section{Mermaid Diagram Source Listings}

The assignment requires editable diagram source files. The report therefore includes Mermaid source
listings for the reconstructed architecture diagrams used as living artifacts.

\subsection{C4 Level 1 --- System Context (Mermaid Source)}
\lstinputlisting{figures/mermaid-src/c4-level-1-system-context.mmd}

\subsection{C4 Level 2 --- Container Diagram (Mermaid Source)}
\lstinputlisting{figures/mermaid-src/c4-level-2-container.mmd}

\subsection{C4 Level 3 --- Backend API Component Diagram (Mermaid Source)}
\lstinputlisting{figures/mermaid-src/c4-level-3-backend-components.mmd}

\subsection{Entity-Relationship Diagram (Mermaid Source)}
\lstinputlisting{figures/mermaid-src/er-data-model.mmd}

\section{Additional Mermaid Source Artifacts Available in the Repository}

The repository also includes supplementary Mermaid sources for dynamic interaction views:

\begin{itemize}
  \item \texttt{final-report/figures/mermaid-src/sequence-document-open-and-sync.mmd}
  \item \texttt{final-report/figures/mermaid-src/sequence-ai-job-flow.mmd}
\end{itemize}

These are maintained as editable assets but are intentionally not expanded into full code listings in the
main report PDF, because the assignment's core diagram requirement is already satisfied through the C4
and ER views included above.
